The brightness and spatial distributions of terrestrial radio sources

### 2013-01-22 articles/Article-RFI-distribution/article-rfi-distribution.tex
Affiliations, etc

\author[A.~R.~Offringa et al.]{
  A.~R.~Offringa$^{1,2}$\thanks{E-mail:
\url{offringa@mso.anu.edu.au}},
A.~G.~de~Bruyn$^{2,3}$,
S.~Zaroubi$^2$%,
%W.~Brouw$^{2,3}$,
%B.~Ciardi$^4$,
%I.~T.~Iliev$^5$,
%\newauthor
%G.~J.~A.~Harker$^6$,
%L.~V.~E.~Koopmans$^2$,
%G.~Mellema$^7$,
%S.~J.~Wijnholds$^3$
\\
$^{1}$Australian National University, RSAA Mt. Stromlo Observatory, Cotter Road, Weston Creek ACT 2611, Australia\\
$^{2}$University of Groningen, Kapteyn Astronomical Institute, PO Box 800, 9700 AV Groningen, The Netherlands\\
$^{3}$ASTRON, PO Box 2, 7990 AA Dwingeloo, The Netherlands\\
%$^{4}$Max-Planck Institute for Astrophysics, Karl-Schwarzschild-Strasse 1, 85748 Garching bei M\"unchen, Germany\\
%$^{5}$University of Sussex, Falmer, Brighton BN1 9QH, UK\\
%$^{6}$Center for Astrophysics and Space Astronomy, University of Colorado Boulder, CO 80309, USA\\
%$^7$Stockholm University, AlbaNova University Center, Stockholm Observatory, SE-106 91 Stockholm, Sweden 
}

@@

\begin{abstract}
Faint undetected sources of radio-frequency interference (RFI) might become visible after long time integration of radio observations if they are stationary and consistently present over time. Thereby, they might obstruct the detection of the weak astronomical signals of interest. This issue is especially important for Epoch of Reionisation (EoR) projects that try to detect the faint redshifted HI signals from the time of the earliest structures in the Universe. We explore the low-frequency RFI situation by studying brightness histograms of visibility data observed with LOFAR, similar to radio source count analyses that are used in cosmology. The histograms shows an RFI distribution that follows a power-law distribution with an estimated exponent around -1.5. With several assumptions, this can be explained with a uniform distribution of terrestrial radio sources whose radiation follows existing propagation models. Extrapolation of the power law would imply that the current LOFAR EoR observations should be severely RFI limited if these sources add up consistently. These observations almost reach the thermal noise and are thought not to be limited by RFI. Therefore, we conclude that it is likely that most RFI sources do not linearly increase the noise, in which case RFI will not prevent an EoR detection with LOFAR.
\end{abstract}

>>>>

\author[A.~R.~Offringa et al.]{
  A.~R.~Offringa$^{1,2,3}$\thanks{E-mail:
\url{offringa@mso.anu.edu.au}},
A.~G.~de~Bruyn$^{3,4}$ and
S.~Zaroubi$^3$%,
%W.~Brouw$^{3,4}$,
%B.~Ciardi$^5$,
%I.~T.~Iliev$^6$,
%\newauthor
%G.~J.~A.~Harker$^7$,
%L.~V.~E.~Koopmans$^3$,
%G.~Mellema$^8$,
%S.~J.~Wijnholds$^4$
\\
$^{1}$Australian National University, RSAA Mt. Stromlo Observatory, Cotter Road, Weston Creek ACT 2611, Australia\\
$^{2}$ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)\\
$^{3}$University of Groningen, Kapteyn Astronomical Institute, PO Box 800, 9700 AV Groningen, The Netherlands\\
$^{4}$ASTRON, PO Box 2, 7990 AA Dwingeloo, The Netherlands\\
%$^{5}$Max-Planck Institute for Astrophysics, Karl-Schwarzschild-Strasse 1, 85748 Garching bei M\"unchen, Germany\\
%$^{6}$University of Sussex, Falmer, Brighton BN1 9QH, UK\\
%$^{7}$Center for Astrophysics and Space Astronomy, University of Colorado Boulder, CO 80309, USA\\
%$^8$Stockholm University, AlbaNova University Center, Stockholm Observatory, SE-106 91 Stockholm, Sweden 
}

@@

\begin{abstract}
Faint undetected sources of radio-frequency interference (RFI) might become visible after long time integration of radio observations if they are stationary and consistently present over time. Thereby, they might obstruct the detection of the weak astronomical signals of interest. This issue is especially important for Epoch of Reionisation (EoR) projects that try to detect the faint redshifted HI signals from the time of the earliest structures in the Universe. We explore the low-frequency RFI situation by studying brightness histograms of visibility data observed with LOFAR, similar to radio source count analyses that are used in cosmology. The histograms show an RFI distribution that follows a power-law distribution with an estimated exponent around -1.5. With several assumptions, this can be explained with a uniform distribution of terrestrial radio sources whose radiation follows existing propagation models. Extrapolation of the power law would imply that the current LOFAR EoR observations should be severely RFI limited if these sources add up consistently. These observations almost reach the thermal noise and are thought not to be limited by RFI. Therefore, we conclude that it is likely that most RFI sources do not linearly increase the noise, in which case RFI will not prevent an EoR detection with LOFAR.
\end{abstract}
